---
title: 'Generalized Linear Models'
subtitle: 'For Over-Dispersed Data'
format: 
  baylor_theme-revealjs:
    author: 'Nathen Byford'
    footer: 'Nathen Byford'
    controls: auto
---

```{r}
#| include: false

library(tidyverse); theme_set(theme_bw())
library(flextable)
```

## Generalized Linear Models (GLMs)

In statistics, a generalized linear model (GLM) is a flexible generalization of ordinary linear regression. The GLM generalizes linear regression by allowing the linear model to be related to the response variable via a link function and by allowing the magnitude of the variance of each measurement to be a function of its predicted value.

## Over-Dispersed Data
- Over-dispersion in general refers to having variance greater than that assumed for the theoretical data model.  
- Over-dispersion can also refer to having a variance that is greater than the mean
    - Similarly equi-dispersion would refer to having variance equal to the mean.

::: {.callout-note}
### Poisson regression

Poisson regression is the most popular method for modeling count data. The Poisson distribution brings with it the assumption of equi-dispersion that is often unsatisfied.
:::


## Candidate Distributions

1. Negative-Binomial
2. Doubble Poisson
3. Generalized Poisson
4. Conway-Maxwell-Poisson (CMP)

## Negative-Binomial

- Parameters: mean: $\mu$, dispersion: $k$[^1]
- Variance: $\mu + \mu^2/k$
    - Function of mean and dispersion parameter
    - Clearly captures over-dispersion.

[^1]: The classic negative-binomial is parameterized as number of success and probability of success, $r$ and $p$.

## Double Poisson

- Parameters: $\lambda$, $\theta$
- variance: $\lambda / \theta$

## Generalized Poisson

- Parameters: $\lambda$, $\alpha$

## Conway-Maxwell Poisson (CMP)

- Parameters: $\lambda$, $\nu$

## Model comparisons

- 2008 Bayesian paper compares the Generalized Poisson distribution,

```{r}
tribble(
  ~Model, ~Poisson, ~NB, ~GP,
  "DIC", 1291.8, 1273.9, 1265.6
) |> 
  flextable()
```

- Bayesian paper compared Poisson, Negative Binomial, and CMP for longitudinal counts using DIC to compare. (Alam et al.)

```{r}
tribble(
  ~Model, ~Poisson, ~NB, ~CMP,
  "DIC", 1362.39, 1350.67, 1348.87
) |>
  flextable()
```

- Another Bayesian paper compared using AIC and shows the following results (Sellers and Shmueli)

```{r}
tribble(
  ~Model, ~CMP, ~Poisson, ~"Neg-Bin",
  "AIC", 5073, 5589, 5077
) |>
  flextable()
```
