---
title: 'Generalized Linear Models'
subtitle: 'For Over-Dispersed Data'
date: ''
format: 
  baylor_theme-revealjs:
    author: 'Nathen Byford'
    footer: 'Nathen Byford'
    controls: auto
---

```{r}
#| include: false

library(tidyverse); theme_set(theme_bw())
library(flextable)
```

## Generalized Linear Models (GLMs)

In statistics, a generalized linear model (GLM) is a flexible generalization of ordinary linear regression. The GLM generalizes linear regression by allowing the linear model to be related to the response variable via a link function and by allowing the magnitude of the variance of each measurement to be a function of its predicted value.

## Over-Dispersed Data
- Over-dispersion in general refers to having variance greater than that assumed for the theoretical data model.  
- Over-dispersion can also refer to having a variance that is greater than the mean
    - Similarly equi-dispersion would refer to having variance equal to the mean.

::: {.callout-note}
### Poisson regression

Poisson regression is the most popular method for modeling count data. The Poisson distribution brings with it the assumption of equi-dispersion that is often unsatisfied.
:::


## Candidate Distributions

- Negative-Binomial
- Poisson-Lindley
- Generalized Poisson
- Conway-Maxwell-Poisson (CMP)

## Negative-Binomial

- Parameters: number of successes: $r$, probability of success: $p$.
- Mean: $r(1-p)/p$
- Variance: $r(1-p)/p^2$

## Poisson-Lindley

- Parameters: $\lambda$, $\theta$

## Generalized Poisson

- Parameters: $\lambda$, $\alpha$

## Conway-Maxwell Poisson (CMP)

- Parameters: $\lambda$, $\nu$

## Model comparisons

Often hard to find papers comparing all  model previously mentioned.

- Bayesian paper compared Poisson, Negative Binomial, and CMP for longitudinal counts using DIC to compare. (Alam et al.)

```{r}
tribble(
    ~Model, ~Poisson, ~NB, ~CMP,
    "DIC", 1362.39, 1350.67, 1348.87
) |>
  flextable()
```

- Another Bayesian paper compared using AIC and shows the following results (Sellers and Shmueli)

```{r}
tribble(
    ~Model, ~CMP, ~Poisson, ~"Neg-Bin",
    "AIC", 5073, 5589, 5077
) |>
  flextable()
```
